---
title: "Closest_Approach_Code"
author: "John Veon"
date: "2022-11-21"
output: html_document
---

##Read in data and packages
```{r}
CAglmer <- read.csv("./ClosestApproachJWMGLMM.csv", header=T,)
library(ggplot2) ### regular plotting with ggplot
library(ggfortify) ### this gives us autoplot
library(MASS) ### this will let us do stepwise regression in an automated fashion
library(ggthemes) # gives us theme_few()
library(lme4)###GLMM
library(lattice)
library(lmerTest)
```

```{r}
getwd()
```


##Variables##
```{r}
CAglmer$Site_NumFactor = as.factor(CAglmer$Site_Num)
CAglmer$Phase_Num = as.factor(CAglmer$Phase_Num)
CAglmer$Distance = CAglmer$Distance
```


###Choosing GLMMs (as opposed to GLMs) to deal with non-independence in response

###little Bimodal, but appears gamma is the fix! How do editors know these things with no data!!!
###Random effect from normal distribution; its odd because hist says 1 has 6, but when I count the data it says 3. I believe this is a histogram glitch.
```{r}
###Assumption check: before model fitting###

##Normality of Response variable
#Should at least be unimodel with a bit of tails to either side
#make transformations as needed
hist(CAglmer$Distance)

##Normality of Random Effect
hist(CAglmer$Site_Num2)
```

```{r}
table(CAglmer$Site_Num)
```


###Variance kind of equal###
###I think this is because birds were super close in phase 1, then moved out in outer phases; So I would say we are okay.
```{r}
##Factors as Predictor variables?
##Homogeneity of variance
#variance should be relatively equal between groups
boxplot(Distance ~ Phase_Num, data = CAglmer)
```



```{r}
CAglmer$Phase_Num2 = relevel(CAglmer$Phase_Num, ref = "2")
CAglmer$Phase_Num = relevel(CAglmer$Phase_Num, ref = "1")
```

```{r}
CAmod1<-glmer(Distance ~ Phase_Num + (1 | Site_NumFactor), data=CAglmer, family = Gamma(link = "log"))
summary(CAmod1)
```

```{r}
confint(CAmod1, method = "Wald")
```



###Phase 2 as ref###
```{r}
##run model
CAmod<-glmer(Distance ~ Phase_Num2 + (1 | Site_NumFactor), data=CAglmer, family = Gamma(link = "log"))
summary(CAmod)
```

```{r}
anova(CAmod)
```

```{r}
confint(CAmod, method = "Wald")
```




###linearity of residuals of response - I dont think we need these; You use these in LMMs to assess linearity, but its assumed here that your link and distribution that were chosen are satisfying this.
```{r}
plot(resid(CAmod),CAglmer$Distance)
```

###linearity of residuals against predictor - I dont think we need these; You use these in LMMs to assess linearity, but its assumed here that your link and distribution that were chosen are satisfying this.
```{r}
plot(resid(CAmod),CAglmer$Phase_Num)
```


###Either way, I think they look good^^^ I traditionally have seen that upward trend in response residuals and then the randomness in fixed effects is what you want.


###Check Normality of residuals of fitted model after using Gamma
```{r}
qqmath(resid(CAmod)) ###fairly normal - good!!!
```

##Checking model variances###
```{r}
plot(CAmod)
plot(resid(CAmod))

#####Check Deviance/df.residuals to see how far from 1 (variances)
##179.9/16 = 11.2 (I'd say chi-square isnt too large)
```

###Check Normality of Random Effect - Normal; passed both tests
```{r}
###Normality of Intercepts
r_int<- lme4::ranef(CAmod)$Site_NumFactor$`(Intercept)`
qqnorm(r_int)
qqline(r_int)
shapiro.test(r_int)

###Normality of Slopes code### --Didnt need this because we did not include random slopes
#r_slope<- lme4::ranef(CAmod)$Site_NumFactor$where_1_is_in((1|Site_NumFactor))
#qqnorm(r_slope)
#qqline(r_slope)
#shapiro.test(r_slope)
```


###Check variances of random effects
###Yay! Variances equal!!!
```{r}
CAglmer$ModelResCA <- residuals(CAmod)
CAglmer$ModelAbsResCA <- abs(CAglmer$ModelResCA)
CAglmer$ModelAbsResSqCA <- CAglmer$ModelAbsResCA^2

levenemodCA <- lm(ModelAbsResSqCA ~ Site_Num, data=CAglmer)
anova(levenemodCA)
```



```{r}
emm <- data.frame(emmeans(CAmod1, ~ Phase_Num, type = "response"))
```

```{r}
#THIS IS THE GOOD ONE#
levels(emm$Phase_Num) <- c("Pre", "During", "Post") 


ClosestApp = ggplot(emm, aes(x= Phase_Num, y=response, fill = Phase_Num))+
  geom_bar(colour = "black", position="dodge", stat="identity", width = 0.4, size = 0.6) + scale_y_continuous(expand = c(0, 0), limits = c(0, 70)) +
  geom_errorbar(aes(ymin=asymp.LCL, ymax = asymp.UCL), width=0.1,size = 0.6, color = "grey0", position = position_dodge(.9)) + 
  ylab("Mean closest approach (m)") + xlab("Phase") +
  scale_fill_manual("legend", values = c("Pre" = "grey85", "During" = "grey40", "Post" = "grey60")) +
  theme_bw() + theme(legend.position = "none", panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_rect(color = "black", size = 1), axis.text = element_text(size = 11, color = "black"), axis.title = element_text(size = 14),)

ClosestApp
```

```{r}
#THIS IS THE GOOD ONE#
levels(emm$Phase_Num) <- c("Pre", "During", "Post") 


ClosestApp = ggplot(emm, aes(x= Phase_Num, y=response, fill = Phase_Num))+
  geom_bar(colour = "black", position="dodge", stat="identity", width = 0.4, size = 0.6) + scale_y_continuous(expand = c(0, 0), limits = c(0, 50)) +
  geom_errorbar(aes(ymin=response-SE, ymax = response+SE), width=0.1,size = 0.6, color = "grey0", position = position_dodge(.9)) + 
  ylab("Mean closest approach (m)") + xlab("Phase") +
  scale_fill_manual("legend", values = c("Pre" = "grey85", "During" = "grey40", "Post" = "grey60")) +
  theme_bw() + theme(legend.position = "none", axis.text.x = element_text(size = 16, color = "black", family = "Arial"), axis.text.y = element_text(size = 16, color = "black", family = "Arial"), axis.title.y = element_text(size = 18, family = "Arial"), axis.title.x = element_text(size = 18, family = "Arial"), legend.title = element_text(size = 16, family = "Arial", face = "bold"), legend.text = element_text(size = 16, color = "black", family = "Arial"), panel.background = element_blank(), panel.border = element_blank(), axis.line = element_line(colour = "black"), panel.grid = element_blank())

ClosestApp
```

